---
title: "HW2"
author: "Florentina Vania"
date: "2024-01-29"
output: html_document
---

```{r setup, include=FALSE}
# install.packages("forecast")
# install.packages("readxl")
library(forecast)
library(readxl)
library(dplyr)
library("tseries")

#rm(list=ls(all=TRUE)) 	# clear data
```

##  Deliverables for HW 2
Loading the dataset. 
Note: The dataset starts from Jan22 .
```{r}
options(scipen = 999) # turn off scientific notation
# Read the data
covid_data <- read_excel("~/Desktop/VANIA/UC Davis 2023/Winter Semester/Topics in Business Analytics I/W2/Class2/2020_Covid_Data.xlsx")

#Covid Data of Particular State
ct_data <- covid_data$VA 

#The dataset contains values starting from Jan22 hence 345
ct_data <- ct_data[1:345]
length(ct_data)
ct_data
```

# 1. Step-by-Step fit different ARIMA (p,d,q) x (P, D, Q) for the confirmed cases. Can you discover a better model than auto.arima?

Convert data to time series object and plot it
```{r}
# Create a time series object using a simple numeric sequence as the time index
ct_ts <- ts(ct_data, start= c(2020,1,22), frequency=345)# Starts from the date Jan 22
ct_ts
## see what would happen if we don't use the whole historical data? e.g. starts from 2020.8, 2023.8,...?

ct_ts
plot.ts(ct_ts) # plot time series first

```
# ARIMA models
Auto ARIMA model
```{r}
#Fit the auto.arima model

auto_fit <- auto.arima(ct_ts)
summary(auto_fit)
# 
# ARIMA(0,2,1) 
# 
# Coefficients:
#           ma1
#       -0.7366
# s.e.   0.0366
# 
# sigma^2 = 153537:  log likelihood = -2534.59
# AIC=5073.17   AICc=5073.21   BIC=5080.85
# 
# Training set error measures:
#                    ME     RMSE      MAE      MPE     MAPE MASE        ACF1
# Training set 48.18329 390.1303 233.9834 1.368118 2.529355  NaN -0.04537713

```

# Step-by-Step
# 2. Select the best model using information criteria (AIC_c, BIC), and present out-of-sample forecasts with prediction intervals
```{r}
## Step 1. Is the time series stationary? 

# Use Augmented Dickey-Fuller Test to test stationary
adf.test(ct_ts)				# if p-value is large (> 0.10), then non stationary
y1 <- diff(ct_ts, differences = 1)			
plot.ts(y1)		# looks stationary visually
adf.test(y1)	# p-value is large
y2 <- diff(ct_ts, differences = 2)	
plot.ts(y2)		# looks stationary visually
adf.test(y2)	# estimated p = 0.01 => small p-value (< 0.10) => so yd is stationary ==> fix d = 2 in ARIMA models to be fitted  -----> here my p value is 0.01


## Step 2. Decide AR(p) or MA(q) or both ARMA(p,q).
Pacf(y2, lag.max = 40) # PACF suggest p=4
Acf(y2, lag.max = 40)	#ACF suggest q=6


# # nearby models Arima, d = 2;p->1,2,3 d-->2 q-->1,2,3
# Create a list to store models and their BIC values
models <- list()
bic_values <- numeric()

# Define the range of p, d, and q
p_values <- 0:1
d_value <- 2
q_values <- 0:2

# Iterate over all combinations of p, d, q
for (p in p_values) {
  for (q in q_values) {
    model <- Arima(ct_ts, order = c(p, d_value, q))
    model_name <- paste("ARIMA(", p, ",", d_value, ",", q, ")", sep = "")
    models[[model_name]] <- model
    bic_values[model_name] <- BIC(model)
  }
}

# Print the summaries and BIC of all models
for (model_name in names(models)) {
  cat("Model:", model_name, "\n")
  print(summary(models[[model_name]]))
  cat("BIC:", bic_values[model_name], "\n\n")
}

# Identify the model with the lowest BIC
best_model_name <- names(which.min(bic_values))
best_model <- models[[best_model_name]]
cat("Model with the lowest BIC:\n")
cat(best_model_name, "\n")
print(summary(best_model))


# predict best_model
y_hat <- predict(best_model)
y_hat

```

# 3. Discover a set of good models in the neighborhood of the best model. Then present the "Consensus Forecast" for COVID cases over the next week (from Sunday to Saturday after the submission date). 
```{r}
# Forecasting for a specified period (e.g., 41 days from the end of the series)
forecasted_values <- forecast(best_model, h=41) 
# Plot the forecast
plot(forecasted_values)
print(forecasted_values)
```
# 4. Your HW score will be based on MAPE (Mean Absolute Percentage Errors) based on actual cases and your forecasts. The smaller the MAPE, the better the model. 
```{r}

# Define the range of p, d, and q
p_values <- 0:1
d_value <- 2
q_values <- 0:2

# Store models
models <- list()

# Iterate over all combinations of p, d, q
for (p in p_values) {
  for (q in q_values) {
    model <- Arima(ct_ts, order = c(p, d_value, q))
    model_name <- paste("ARIMA(", p, ",", d_value, ",", q, ")", sep = "")
    models[[model_name]] <- model
  }
}

# Print the summary of all models
for (model_name in names(models)) {
  cat("Model:", model_name, "\n")
  print(summary(models[[model_name]]))
  cat("\n")
}



# ARIMA(1,2,1): MAPE = 3.099501
# ARIMA(1,2,2): MAPE = 3.279466
# ARIMA(1,2,3): MAPE = 3.271208
# ARIMA(2,2,1): MAPE = 3.310667
# ARIMA(2,2,2): MAPE = 3.321913
# ARIMA(2,2,3): MAPE = 3.25787
# ARIMA(3,2,1): MAPE = 3.405724
# ARIMA(3,2,2): MAPE = 3.384065
# ARIMA(3,2,3): MAPE = 3.300773
# Based on these values, the model with the lowest MAPE is:
# 
# ARIMA(1,2,1) with a MAPE of 3.099501
# Therefore, the ARIMA(1,2,1) model is the best in terms of forecasting accuracy according to the MAPE criterion among the models listed. Remember, a lower MAPE value indicates a better fit of the model to the data

# Model: ARIMA(1,2,1) 
# Series: ct_ts 
# ARIMA(1,2,1) 
# 
# Coefficients:
#           ar1      ma1
#       -0.0544  -0.7100
# s.e.   0.0762   0.0575
# 
# sigma^2 = 153753:  log likelihood = -2534.33
# AIC=5074.65   AICc=5074.72   BIC=5086.16
# 
# Training set error measures:
#                    ME     RMSE      MAE      MPE     MAPE MASE        ACF1
# Training set 46.63337 389.8337 233.8804 1.356576 2.523652  NaN -0.01368202
```

```{r}
# Fit the best ARIMA model based on lowest MAPE -substitute order which the combination that gave the lowest MAPE

best_model_MAPE <- Arima(ct_ts, order = c(1,2,1))
# Forecasting for a specified period (e.g., 41 days from the end of the series)
forecasted_values <- forecast(best_model_MAPE, h=41) 
# Plot the forecast
plot(forecasted_values)
print(forecasted_values)
print(forecasted_values$mean)

#Result of Vania Forecast State = VA
# Time Series:
# Start = c(2021, 1) 
# End = c(2021, 41) 
# Frequency = 345 
#  [1] 359151.1 363578.7 368003.9 372429.2 376854.5 381279.9 385705.2 390130.5 394555.9
# [10] 398981.2 403406.5 407831.9 412257.2 416682.5 421107.8 425533.2 429958.5 434383.8
# [19] 438809.2 443234.5 447659.8 452085.2 456510.5 460935.8 465361.1 469786.5 474211.8
# [28] 478637.1 483062.5 487487.8 491913.1 496338.5 500763.8 505189.1 509614.5 514039.8
# [37] 518465.1 522890.4 527315.8 531741.1 536166.4
```